### 2013-09-18 exact_learning/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

\section{Learning Structure Support Vector Machines with Approximate Inference}
In this chapter, we use the 1-slack cutting plane algorithm, as we found this
algorithm to be most suitable for out approach. For reference, the algorithm is outlined
\Secref{one_slac}, Algorithm~\ref{alg_one_slack}. The inference algorithm $I$ is called
in line~\ref{get_cutting_plane}. We investigate algorithms $I$ (often called separation
oracles in the context) that do not yield the exact maximum here.
There are two groups of inference procedures, as identified in
\citet{finley2008training}: under-generating and over-generating approaches.
An under-generating approach satisfies $I(x^i, y^i, \theta) \in
\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane}
of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
hand, does solve the loss-augmented prediction in line~\ref{get_cutting_plane}
exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.
%We elaborate on
%the properties of under-generating and over-generating inference procedures in
%Section~\ref{bounds}. %TODO Filler sentence? wat?

>>>>

\section{Learning SSVMs with Approximate Inference}
In this chapter, we use the 1-slack cutting plane algorithm, as we found this
algorithm to be most suitable for out approach. For reference, the algorithm is outlined
\Secref{one_slac}, Algorithm~\ref{alg_one_slack}. The inference algorithm $I$ is called
in line~\ref{get_cutting_plane}. We investigate algorithms $I$ (often called separation
oracles in the context) that do not yield the exact maximum here.
There are two groups of inference procedures, as identified in
\citet{finley2008training}: under-generating and over-generating approaches.
An under-generating approach satisfies $I(x^i, y^i, \theta) \in
\mathcal{Y}$, but does not guarantee maximality in line~\ref{get_cutting_plane}
of Algorithm~\ref{alg_cutting_plane}. An over-generating approach on the other
hand, does solve the loss-augmented prediction in line~\ref{get_cutting_plane}
exactly, but for a larger set $\hat{\mathcal{Y}} \supset \mathcal{Y}$, meaning
that possibly $I(x^i, y^i, \theta) \notin \mathcal{Y}$.
%We elaborate on
%the properties of under-generating and over-generating inference procedures in
%Section~\ref{bounds}. %TODO Filler sentence? wat?

### 2013-09-18 intro.tex
more fixes by hannes, working on introduction, adding xkcd image

\chapter{Introduction}
The task of semantic segmentation is defined as predicting a semantic class label
for each pixel in a natural image. Predicting semantic labels on such a fine scale
constitutes an important step towards general scene understanding.
%
Applications of scene understanding include automatic image interpretation for retrieval,
autonomous driving and mobile robotics.
%
We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent the input.
%
%The main part of this thesis investigates the tasks of semantic segmentation and object class segmentation in natural images.
The main part of this thesis investigates the use of structured learning
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities lately.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.  We hoe this thesis will further the understanding
of structured learning methods, give practical guidance and provide
implementations to further the use of structured prediction methods in computer
vision.

@@

We will focus on the use of conditional random fields, which have shown very
promising.  Using the paradigm of structural support vector machines, it is
possible to learn conditional random fields to directly minimize a loss of
interest. In particular, conditional random fields allow to combine different
cues, possibly produced by other approaches, in a principled manner.
%
Why semantic segmentation? Difference object / stuff centric? -> difference to Alex
Structured approaches
%

\section{List of Contributions}
\begin{itemize}
\item Introduction of general clustering algorithm that improves upon widely used
    approaches from the literature.
\item Demonstration of an algorithm for  weakly supervised object class segmentation.
\item Providing a general efficient implementation for structured prediction.
\item Analysis of max-margin learning algorithms with exact or approximate inference in different applications.
\item Demonstrating that exact learning for semantic segmentation and object class segmentation is possible in loopy graphs.
%\item Shows importance of edge features.
\end{itemize}

@@

We will then introducing a mostly unstructured
approach to semi-supervised learning for object class segmentation in
Chaper~\ref{ch:semi_supervised}, investigating the problem of annotating
training data for semantic segmentation. The central topic of this thesis is
introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
software library for implementing structured learning and prediction
algorithms.
Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
used structured prediction algorithms, focusing on which algorithm is best to
use in a given situation. In particular, we investigate learning behavior for
semantic segmentation on several datasets.

@@

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structurd_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spacial relations, and improve
upon the state-of-the-art in the NYU V2 datasets that we consider.

\section{Publications}
The main material of this thesis has either been published in conference
proceedings or has been submitted to conferences or journals. We now list the
relevant publications.
\begin{description}
    \item[Chapter 2] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C. M\"uller, Sebastian Nowozin and Christoph H. Lampert. Published in the proceedings of the German Conference on Pattern Recognition.
    \item[Chapter 3] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven Behnke. Published in the proceedings of the IARP Worshop on Partially Supervised Learning.
    \item[Chapter 4] \emph{PyStruct - Structured Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of Machine Learning Research, Open Source Software track.
    \item[Chapter 5] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}. Andreas C. M\"uller and Sven Behnke. Submitted to the International Conference on Computer Vision Theory and Applications.
    \item[Chapter 6] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}. Andreas C. M\"uller and Sven Behnke. Submitted to the International Conference on Robotics and Automation.
\end{description}

>>>>

\chapter{Introduction}
\vspace{1cm}
\hfill%
\begin{minipage}{.4\linewidth}
\includegraphics[width=\linewidth]{python_soul}%
\flushright%
Randall Munroe, \emph{xkcd.com}
\end{minipage}%
\\[2cm]
In computer vision research, the goal is to automatically extract information from an image.
In particular discerning semantic information, that is interpreting an image, is a prominent
research topic.
While much process has been made in recent years, computer vision systems still lack behind
humans in most tasks that require semantic information. Many semantic tasks can be formulated
in terms of \emph{semantic classes}, that is categories of parts, objects or scenes.
Examples include answering questions such as ``Is this a picture of a beach?'', ``How many
cars are in this image'' or even ``What objects lie on the table?''.
These questions illustrate a range of possible tasks involving semantic categories,
from classifying images of single objects, over localization and counting of object classes,
to parsing a scene fully into objects and object classes, together with their relations.
%
While humans can distinguish tens of thousands of object classes, and have little trouble
in interpreting complex scenes, current methods are often restricted to a much smaller number
of classes. While some recent approaches are able to classify images into a thousand 
or even ten thousand classes, results are still somewhat mixed.
In particular, when refining the task to localization, and allowing multiple objects within
a scene, few approaches have been evaluated on more than about twenty classes.

In this work, we will investigate the tasks of semantic segmentation and object
class segmentation.  These ask for a semantic annotation of each pixel with one
of a set of known classes. We choose this task in particular for several reasons:
\begin{itemize}
    \item Pixel-level annotation provides highly detailed information about the scene.
    \item Joint estimation of multiple classes allows for the use of \emph{context}.
    \item In contrast to category-agnostic segmentation approaches, object class segmentation
        has an unambiguous true labeling.
    \item A variety of annotated datasets is publicly available.
\end{itemize}
%
Applications of semantic segmentation and scene understanding include automatic
image interpretation for retrieval, autonomous driving and mobile robotics.
While these are already important applications, we expect that with the
abundance of camera data, and the proliferation of mobile computing, even
more applications will gain importance soon.

We will in the following distinguish between the task of semantic segmentation,
which is usually distinguishing unstructured ``stuff'' classes such as road and
grass, and object class segmentation, which denotes the segmentation of very
structured classes, such as cars, planes and people.
We consider four different datasets in this thesis: the object class segmentation
datasets Graz02 and Pascal VOC 2010, and the semantic segmentation datasets MSRC-21,
and NYU V2.
%
Both tasks have the same ultimate goal of parsing, and therefore understanding, images
in terms of semantic classes, but usually employ  different mechanisms to represent the input.

One of the bottlenecks in learning object class segmentation and semantic
segmentation is the availability of training data.  While unlabeled image data,
and even data with semantic
``tags'' is available in practically unlimited quantities, semantic annotation
on pixel level is only available through laborious manual annotation. 
Chapter %FIXME
introduces an approach to cope with this shortcoming of object class segmentation approaches
by introducing a method to find segmentations automatically from image level annotations.

The main part of this thesis investigates the use of structured learning
algorithms to the task of semantic image segmentation. Both topics have receive
much attention in the computer vision and machine learning communities lately.
Unfortunately, learning structured prediction in computer vision applications
is still little understood.

We will focus on the use of conditional random fields~(CRFs), which have shown
great promise for computer vision applications. Using the paradigm of
structural support vector machines, it is possible to learn conditional random
fields to directly minimize a loss of interest. In particular, conditional
random fields allow to combine different cues, possibly produced by other
approaches, in a principled manner.

There have been several previous studies on learning structural support vector machines,
and learning for conditional random fields. The impact of approximate inference was first
investigated by \citet{finley2008training}, applying structural support vector machines
to multi-label data. Later, different works investigated how to combine approximate inference and
learning in a single framework. \citet{meshi2010learning, komodakis2011efficient}, and \citet{hazan2010primal} approached
the problem using duality, and formulate learning and inference as a joint optimization problem.
\citet{stoyanov2011empirical}, and later \citet{jancsarylearning} and  \citet{krahenbuhlparameter} formulate learning
structured prediction as optimizing an \emph{prediction engine}, that takes into account all aspects
of the model, in particular the inference algorithm used.
In this work, on the other hand, we follow the well-established algorithms for learning structural
support vector machines, and investigate how we can use the available inference algorithms to
obtain good results within a reasonable time-frame.

\citet{nowozin2010parameter} provide a detailed evaluation of different aspects of learning object class
segmentation, that is somewhat orthogonal to this work. Their work considers the choice of features,
number superpixels and pairwise potentials for conditional maximum likelihood learning of tree-structured CRFs.
\citet{nowozin2010parameter} also compare conditional maximum likelihood learning with maximum margin learning,
finding little difference in accuracy.
We focus our work on the more popular neighborhood graphs, which do not allow for easy inference.
Therefore, conditional maximum likelihood learning would be much harder in our setting. For our studies,
we fix the considered feature set, and vary the maximum margin learning algorithm. Despite
this seemingly limited setup, we can make out large differences in the applicability of
some of the commonly used methods. For comparison, we also evaluate the learning algorithms
on chain CRFs, where inference is simple.

%We hoe this thesis will further the understanding
%of structured learning methods, give practical guidance and provide
%implementations to further the use of structured prediction methods in computer
%vision.

@@

%
%Structured approaches
%

\section{List of Contributions}
This thesis contains the following contributions:
\begin{itemize}
\item Introduction of general clustering algorithm that improves upon widely used
    approaches from the literature.
\item Demonstration of an algorithm for  weakly supervised object class segmentation.
\item Providing a general efficient implementation for structured prediction.
\item Analysis of max-margin learning algorithms with exact or approximate inference in different applications.
\item Demonstrating that exact learning for semantic segmentation and object class segmentation is possible in loopy graphs.
%\item Shows importance of edge features.
\end{itemize}

@@

We will then introducing a mostly unstructured
approach to semi-supervised learning for object class segmentation in
Chapter~\ref{ch:semi_supervised}, investigating the problem of annotating
training data for semantic segmentation. The central topic of this thesis is
introduced in Chapter~\ref{ch:structured_pystruct}, which also introduces our
software library for implementing structured learning and prediction
algorithms.
Chapter~\ref{ch:comparision} gives a quantitative comparison of the most widely
used structured prediction algorithms, focusing on which algorithm is best to
use in a given situation. In particular, we investigate learning behavior for
semantic segmentation on several datasets.

@@

Finally, Chapter~\ref{ch:nyu} applies the methods developed in
Chapter~\ref{ch:structured_pystruct} and Chapter~\ref{ch:exact_learning} to the
problem of semantic annotation with structure classes in RGB-D data. We
demonstrate that we are able to learn meaningful spacial relations, and improve
upon the state-of-the-art in the NYU V2 datasets that we consider.

\section{Publications}
The main material of this thesis has either been published in conference
proceedings or has been submitted to conferences or journals. We now list the
relevant publications.
\begin{description}
    \item[Chapter~\ref{ch:itm}] \emph{Information Theoretic Clustering using Minimum Spanning Trees} Andreas C. M\"uller, Sebastian Nowozin and Christoph H. Lampert. Published in the proceedings of the German Conference on Pattern Recognition.
    \item[Chapter~\ref{ch:semi_supervised}] \emph{Multi-Instance Methods for Partially Supervised Image Segmentation} Andreas C. M\"uller and Sven Behnke. Published in the proceedings of the IARP Worshop on Partially Supervised Learning.
    \item[Chapter~\ref{ch:structured_pystruct}] \emph{PyStruct - Structured Prediction in Python} Andreas C. M\"uller and Sven Behnke. Submitted to the Journal of Machine Learning Research, Open Source Software track.
    \item[Chapter~\ref{ch:exact_learning}] \emph{Learning a Loopy Model for Semantic Segmentation Exactly}. Andreas C. M\"uller and Sven Behnke. Submitted to the International Conference on Computer Vision Theory and Applications.
    \item[Chapter~\ref{ch:nyu}] \emph{Learning Depth-Sensitive Conditional Random Fields for Semantic Segmentation}. Andreas C. M\"uller and Sven Behnke. Submitted to the International Conference on Robotics and Automation.
\end{description}

### 2013-09-18 itm/paper.tex
more fixes by hannes, working on introduction, adding xkcd image

\chapter{Information Theoretic Clustering}

@@

Clustering algorithms based on minimum spanning trees have been studied early
on in the statistics community, due to their efficiency.  One of the earliest
methods is single-link agglomerative clustering~\citep{gower1969minimum}.
Single-link agglomerative clustering can be understood as a minimum spanning
tree-based approach in which the largest edge is removed until the desired
number of components is reached.  \citet{zahn1971graph} refines this criterion
by cutting edges that are longer than other edges in the vicinity. This
approach requires tuning several constants by hand. More recently,
\citet{grygorash2006minimum} proposed a hierarchical MST-based clustering
approach that iteratively cuts edges, merges points in the resulting
components, and rebuilds the spanning tree.
We limit our discussion to the most widely used algorithm from~\citep{gower1969minimum}.
%
\section{Clustering Using Nonparametric Entropy-Estimates}
In general, the goal of clustering can be formulated as follows: 
given a finite collection of samples $\mathbf{x} = (x_1, \dotsc, x_n)$, we want to 
assign cluster-memberships $\mathbf{y} = (y_1, \dotsc, y_n ), y_i \in \{1, \dotsc k\}$ to these samples.
%
We adopt the viewpoint of information theoretic clustering of \citet{gokcay2002information}, 
where the $x_i$ are considered i.i.d{.} samples from a distribution $p(X)$, and the 
$y_i$ are found such that the mutual information $I(X, Y)$ between the distribution $p(X)$ and the
assigned labels $p(Y)$ is maximized.
We can rewrite this objective as 
\begin{align}
         I(X, Y)
        &= D_{\text{KL}}(p(X, Y) \parallel p(X) p(Y)) = H(X) - \sum_{y=1}^k p(Y{=}y) H(X \given Y {=} y)
\end{align}
where
\begin{itemize}
\item $\displaystyle D_\text{KL}\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
\item $\displaystyle H(X)\, = \int_\mathcal{X} p(X) \ln(p(X))\,dX$ is the differential entropy, and
\item $\displaystyle H(X \given Y{=}y)\, = \int_\mathcal{X} p(X \given Y {=}y) \ln(p(X \given Y {=} y))\,dX$ is the conditional differential entropy.
    \end{itemize}
Expressing the mutual information in terms of the entropy is convenient, since
the objective then decomposes over the values of $Y$. 
%
Additionally, $H(X)$ is independent of the distribution of $Y$ and therefore
does not influence the search over $\mathbf{y}$.

>>>>

\chapter{Information Theoretic Clustering}\label{ch:itm}

@@

Clustering algorithms based on minimum spanning trees have been studied early
on in the statistics community, due to their efficiency.  One of the earliest
methods is single-link agglomerative clustering~\citep{gower1969minimum}.
Single-link agglomerative clustering can be understood as a minimum spanning
tree-based approach in which the largest edge is removed until the desired
number of components is reached.  \citet{zahn1971graph} refines this criterion
by cutting edges that are longer than other edges in the vicinity. This
approach requires tuning several constants by hand. More recently,
\citet{grygorash2006minimum} proposed a hierarchical MST-based clustering
approach that iteratively cuts edges, merges points in the resulting
components, and rebuilds the spanning tree.
We limit our discussion to the most widely used algorithm from~\citep{gower1969minimum}.
%
\section{Clustering Using Non-Parametric Entropy Estimates}
In general, the goal of clustering can be formulated as follows: 
given a finite collection of samples $\mathbf{x} = (x_1, \dotsc, x_n)$, we want to 
assign cluster-memberships $\mathbf{y} = (y_1, \dotsc, y_n ), y_i \in \{1, \dotsc k\}$ to these samples.
%
We adopt the viewpoint of information theoretic clustering of \citet{gokcay2002information}, 
where the $x_i$ are considered i.i.d{.} samples from a distribution $p(X)$, and the 
$y_i$ are found such that the mutual information $I(X, Y)$ between the distribution $p(X)$ and the
assigned labels $p(Y)$ is maximized.
We can rewrite this objective as 
\begin{align}
         I(X, Y)
        &= D_{\text{KL}}(p(X, Y) \parallel p(X) p(Y)) = H(X) - \sum_{y=1}^k p(Y{=}y) H(X \given Y {=} y)
\end{align}
where
\begin{itemize}
\item $\displaystyle D_\text{KL}\, = \int_\mathcal{X} p(X) \ln\left(\frac{p(X)}{q(X)}\right)\,dX$ is the Kullback-Leibler divergence,
\item $\displaystyle H(X)\, = \int_\mathcal{X} p(X) \ln(p(X))\,dX$ is the differential entropy, and
\item $\displaystyle H(X \given Y{=}y)\, = \int_\mathcal{X} p(X \given Y {=}y) \ln(p(X \given Y {=} y))\,dX$ is the conditional differential entropy.
    \end{itemize}
Expressing the mutual information in terms of the entropy is convenient, since
the objective then decomposes over the values of $Y$. 
%
Additionally, $H(X)$ is independent of the distribution of $Y$ and therefore
does not influence the search over $\mathbf{y}$.

### 2013-09-18 main.tex
more fixes by hannes, working on introduction, adding xkcd image

\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{refstyle}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
%\usepackage{pslatex}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{tabu}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{mdwlist}
\usepackage{url}
\usepackage[noend]{algpseudocode}
\usepackage[small]{caption}
\usepackage{setspace}
\usepackage{color}
\usepackage{minted}

@@

%\newtheorem{corollary}{Corollary}
%\newtheorem{corollarya}{Corollary}
%\newtheorem{lemma}{Lemma}
%\newtheorem{theorem}{Theorem}
%\newtheorem{theorema}{Theorem}

@@

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction For Semantic Image Segmentation}}\\[5mm]
Dissertation\\%[3mm]
zur\\%[3mm]
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\%[3mm]
der\\%[3mm]
Mathematisch-Naturwissenschaftlichen Fakult\"at\\%[3mm]
der\\%[3mm]
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\%[3mm]
vorgelegt von\\%[5mm]
{\LARGE Andreas Christian M\"uller}\\%[5mm]
aus\\%[3mm]
Offenbach am Main\\%[10mm]
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des machinellen Sehens.
Diese Arbeit untersucht drei m\"oglichen Ansi\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
Baustein in vielen Applikationen des Machinellen Sehens, insbesondere in der
automatischen Segmentierung.

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
mit Hilfe von Supervision in Form von Klassen-Pr\"asenz auf Bildern in der Lage
ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.

Der dritte und letzte Teil er Arbeit untersucht einen der am weitest
verbreiteten Ans\"atze zu semantischen Segmentierung und
Objektklassensegmentierung, n\"amlich Conditional Random Fields verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Mit den vorgestellten Methoden bringen den Stand der Kunst auf
auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

Wir stellen ausserdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisation der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

\chapter*{Abstract}
Automatic segmentation and recognition of semantic classes in natural images is
an important open problem in computer vision.
This work investigates three different approaches to recognition: without supervision,
with supervision on level of images, and with supervision on the level of pixels.
The thesis comprises three parts.

>>>>

\usepackage{hyperref}
\usepackage{epsfig}
\usepackage{subfigure}
\usepackage{calc}
\usepackage{refstyle}
\usepackage{amssymb}
\usepackage{amstext}
\usepackage{amsmath}
\usepackage{amsthm}
\usepackage{multicol}
%\usepackage{pslatex}
\usepackage{microtype}
\usepackage{natbib}
\usepackage{booktabs}
\usepackage{tabularx}
\usepackage{tabu}
\usepackage{graphicx}
\usepackage{algorithm}
\usepackage{mdwlist}
\usepackage{url}
\usepackage[noend]{algpseudocode}
\usepackage[small]{caption}
\usepackage{setspace}
\usepackage{color}
\usepackage{minted}
\usepackage[german,english]{babel}

@@

@@

\begin{center}
\Large\sffamily
\vspace*{1cm}
\textbf{{\huge Learning Structured Prediction for Semantic Image Segmentation}}\\[5mm]
Dissertation\\
zur\\
Erlangung des Doktorgrades (Dr.\ rer.\ nat.)\\
der\\
Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der\\
Rheinischen Friedrich-Wilhelm Universit\"at Bonn\\
vorgelegt von\\
{\LARGE Andreas Christian M\"uller}\\
aus\\
Offenbach am Main\\
Bonn, September 2013
\end{center}
\end{titlepage}
%
%\thispagestyle{empty}
\clearpage
\pagenumbering{Roman}
\setcounter{page}{2}
\vspace*{5cm}
\Large\sffamily
\begin{center}
Angefertigt mit Genehmigung\\
der Mathematisch-Naturwissenschaftlichen Fakult\"at\\
der Rheinischen Friedrich-Wilhelms-Universit\"at Bonn\\
\end{center}
1. Gutachter Prof. Dr. Sven Behnke\\
2. Gutachter Prof. Dr. Reinhard Klein\\
Tag der Promotion:\\
Erscheinungsjahr: 2013\\
\end{spacing}
\tableofcontents
%\listoffigures
%\listoftables
\chapter*{Zusammenfassung}
\enlargethispage{10mm}
\selectlanguage{german}
Automatische Segmentierung und Erkennung von semantischen Klassen in
nat\"urlichen Bildern ist ein wichtiges offenes Problem des maschinellen Sehens.
Diese Arbeit untersucht drei m\"oglichen Ans\"atzen der Erkennung:
ohne \"Uberwachung, mit \"Uberwachung auf Ebene von Bildern und mit \"Uberwachung auf Ebene
von Pixeln. Diese Arbeit setzt sich aus drei Teilen zusammen.

Im ersten Teil der Arbeit schlagen wir einen Clustering-Algorithmus vor,
der eine neuartige, informationstheoretische Zielfunktion optimiert. Wir zeigen, dass
der vorgestellte Algorithmus \"ublichen Standardverfahren aus der Literatur gegen\"uber
klare Vorteile auf vielen verschiedenen Datens\"atzen hat. Clustering ist ein wichtiger
Baustein in vielen Applikationen des machinellen Sehens, insbesondere in der
automatischen Segmentierung.

Der zweite Teil dieser Arbeit stellt ein Verfahren zur automatischen
Segmentierung und Erkennung von Objektklassen in nat\"urlichen Bildern vor, der
mit Hilfe von Supervision in Form von Klassen-Vorkommen auf Bildern in der Lage
ist ein Segmentierungsmodell zu Lernen. Nach unserem Wissen ist dies der erste
Ansatz dieser Art, der sich speziell auf komplexe Objektklassen bezieht.

Der dritte und letzte Teil der Arbeit untersucht einen der am weitesten
verbreiteten Ans\"atze zu semantischen Segmentierung und
Objektklassensegmentierung, Conditional Random Fields, verbunden mit Verfahren
der strukturierten Vorhersage.
%Die Vorhersage f\"ur Modelle zur Bild-Markierung
%mittels Conditional Random Fields ist f\"ur gew\"ohnlich nur approiximativ
%moeglich, bedingt d\"urch das Vorhandensein von Kreisen in den zugrunde
%liegenden Nachbarschaftsgraphen.  Aus diesem Grund, und trotz der Beliebtheit
%von Conditional Random Field Verfahren, ist das strukturierte Lernen f\"ur
%solche Anwendungen schlecht verstanden.
%
Wir untersuchen verschiedene Lernalgorithmen des struktuierten Lernens, insbesondere
im Zusammenhang mit approximativer Inferenz. Wir zeigen, dass es m\"oglich ist trotz des Vorhandenseins von
Kreisen in den Nachbarschaftsgraphen exakte strukturierte Modelle zur Bildsegmentierung zu lernen.
Mit den vorgestellten Methoden bringen den Stand der Kunst
auf zwei komplexen Datens\"atzen zur semantischen Segmentierung voran, dem MSRC-21 Datensatz
von RGB-Bildern und dem NYU V2 Datensatz von RGB-D Bildern von Innenraum-Szenen.

Wir stellen au{\ss}erdem eine Software-Bibliothek vor, die es erlaubt einen
weitreichenden Vergleich der besten Lernverfahren f\"ur strukturiertes Lernen
durchzuf\"uhren.
Unsere Studie erlaubt uns eine Charakterisierung der betrachteten Algorithmen
in einer Reihe von Anwendungen, insbesondere der semantischen Segmentierung und
Objektklassensegmentierung.

\selectlanguage{english}
\chapter*{Abstract}
Automatic segmentation and recognition of semantic classes in natural images is
an important open problem in computer vision.
This work investigates three different approaches to recognition: without supervision,
with supervision on level of images, and with supervision on the level of pixels.
The thesis comprises three parts.
